Installation of Libraries

```{r}
library('ggplot2') # visualization
library('car') # visualization
library('scales') # visualization
library('AER') #Coefficients
require("tidyr")
library('corrplot')
#library('caret')
library('purrr')
library('coefplot')
library('psych')
library('MASS')
#library('leaflet.extras')
library("PerformanceAnalytics")
library('GPArotation')
library('MVN')
library('psych')
library('MASS')
library('psy')
library('corpcor')
library('nnet')
library('plyr')
library('car')
library("e1071")
library('ggcorrplot')
library('mlogit')  # for multiple class logistic regression
# library('InformationValue')
library('rpart.plot')
library('caTools')
library('ggplot2')
library('RColorBrewer')
# library('data.table')
# library('ROCR')
# library('maptree')
# library('tree')
library('dummies') # for converting categorical into dummy one
library('caret')
library('pscl') ## for  McFadden R2
library('randomForest')
library('StatMeasures')
library('sqldf')
library('purrr')
library('tidyr')
#library('caret')
library('ggplot2')
library('gains')
library('lubridate')
library('dummies')
library('glmnet')
library('gbm')
library('VIM')  ### This is for knn
library('DMwR')  ## package for SMOTE
library('rpart')
```
Data Loading

```{r}
setwd("C:/Users/Karthik/Documents/Lakshmi - Back up docs/Learning/Data Analytics/BABI/Sessions/S15/SCLA Assgn")
Data<-read.csv("SCLA_Inventory (1).csv")

```
Basic Data Analysis
```{r}
nrow(Data)
View(Data)
str(Data)
summary(Data)
```
Observation: The dataset is about inventory of a firm. The data consists of 7,853 rows with 8 columns. The data spans from 2007 to 2010 timeframe where the business seems to be steadily growing. The objective of the exercise is to find the optimal ship mode for the order placed by the customers.
```{r}
head(Data, 5)

tail(Data, 5)

dim(Data)

colnames(Data)

```
Converting data types of the variables
Converting the Order.Date to month day year format.
```{r}
Data$Order.Date <- mdy(Data$Order.Date)
View(Data)
```
Exploratory data analysis & interpretations

```{r}
levels(as.factor(Data$Ship.Mode))

Data$Ship.Mode <- factor(Data$Ship.Mode, levels= c("Delivery Truck", "Express Air", "Regular Air"))

mean(Data$Ship.Mode == 'Delivery Truck')
mean(Data$Ship.Mode == 'Express Air')
mean(Data$Ship.Mode == 'Regular Air')


```
```{r}
p1 <- ggplot(Data, aes(x= Ship.Mode, y=Sales, fill=Sales)) + geom_boxplot()  
p2 <- ggplot(Data, aes(x=Ship.Mode, y=Order.Quantity, fill=Order.Quantity)) + geom_boxplot()  

grid.arrange(p1, p2, ncol=2)

```
Observation: The above boxplots Ship.Mode Vs Sales & Order.Quantity shows following observations
1.	Sales numbers are higher when using Delivery_Truck as the shipping mode. It may be due to Delivery truck can be used to deliver large quantities of the orders
2.	The median order quantity for all the 3 modes of Ship_mode. It suggests that irrespective of any shipping mode the median order quantity remains same.

Orders by Ship_mode
```{r}
ship_mode <- aggregate(Data$Ship.Mode, by=list(Data$Ship.Mode), FUN=length)
colnames(ship_mode)<-c("ship_mode","Count")
ship_mode[order(-ship_mode$Count),]
```
Observation: Regular Air Shipping mode has the highest order followed by Delivery Truck. Express Air Shipping mode has lowest orders. It may be due to higher charges levied on Express modes.

Orders by Product Sub-Category
```{r}
Product_Sub_Category <- aggregate(Data$Product.Sub.Category, by=list(Data$Product.Sub.Category), FUN=length)
colnames(Product_Sub_Category)<-c("Product_Sub_Category","Count")
head(Product_Sub_Category[order(-Product_Sub_Category$Count),],5)
```
Orders by Product Name
```{r}
Product_Name <- aggregate(Data$Product.Name, by=list(Data$Product.Name), FUN=length)
colnames(Product_Name)<-c("Product_Name","Count")
head(Product_Name[order(-Product_Name$Count),],5)
```
Sales by Product Category

```{r}
sales_prod_cat <-aggregate(Data$Sales, by=list(Category= Data$Product.Sub.Category), FUN=sum)
colnames(sales_prod_cat)<-c("Product_sub_category","Sales")
head(sales_prod_cat[order(-sales_prod_cat$Sales),],5)
```
Sales by Ship_mode

```{r}
sales_ship_mode <-aggregate(Data$Sales, by=list(Category= Data$Ship.Mode), FUN=sum)
colnames(sales_ship_mode)<-c("Ship_mode","Sales")
head(sales_ship_mode[order(-sales_ship_mode$Sales),],5)
```
Total Sales Trend
Extract days from order dat

```{r}
order_dates = Data$Order.Date

head(order_dates)

order_days <- weekdays(order_dates) 

head(order_days)

#Extract the year from the data
order_years <- substring(order_dates ,1,4)
head(order_years)

sales_days <- data.frame(order_days,Data$Sales,order_dates,order_years)
head(sales_days)

#Edit level order
sales_days$order_days <- factor(sales_days$order_days, levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday"))

```
```{r}
# Sales
ggplot(data=sales_days, aes(x= order_days, y=Data$Sales))+
  geom_bar(stat="identity",show.legend = TRUE,fill="blue")+
  ggtitle("Total sales :Monday to Sunday")+
  facet_wrap(~order_years,ncol=2)+ylab("Sales")

# Order Quantity
ggplot(data=sales_days, aes(x= order_days, y=Data$Order.Quantity))+
  geom_bar(stat="identity",show.legend = TRUE,fill="blue")+
  ggtitle("Order Quantity :Monday to Sunday")+
  facet_wrap(~order_years,ncol=2)+ylab("Order_Quantity")
```
```{r}
boxplot(Data$Order.Quantity ~Data$Ship.Mode, main = "Ship mode vs Order Quantity")
boxplot(Data$Sales ~Data$Ship.Mode, main ="Ship mode vs Sales")

boxplot(Data$Order.Quantity ~Data$Product.Container+Data$Ship.Mode, main = "Product container+Ship mode vs Order Quantity")
boxplot(Data$Sales ~Data$Product.Container+Data$Ship.Mode, main ="Product container+Ship mode vs Sales")

```
```{r}
hist(Data$Order.Quantity, col = 'blue')
hist(Data$Sales, col = 'green')

```
```{r}

#outlier function
outlierKD <- function(dt, var) {
  var_name <- eval(substitute(var),eval(dt))
  na1 <- sum(is.na(var_name))
  m1 <- mean(var_name, na.rm = T)
  par(mfrow=c(2, 2), oma=c(0,0,3,0))
  boxplot(var_name, main="With outliers")
  hist(var_name, main="With outliers", xlab=NA, ylab=NA)
  outlier <- boxplot.stats(var_name)$out
  mo <- mean(outlier)
  var_name <- ifelse(var_name %in% outlier, NA, var_name)
  boxplot(var_name, main="Without outliers")
  hist(var_name, main="Without outliers", xlab=NA, ylab=NA)
  title("Outlier Check", outer=TRUE)
  na2 <- sum(is.na(var_name))
  cat("Outliers identified:", na2 - na1, "n")
  cat("Propotion (%) of outliers:", round((na2 - na1) / sum(!is.na(var_name))*100, 1), "n")
  cat("Mean of the outliers:", round(mo, 2), "n")
  m2 <- mean(var_name, na.rm = T)
  cat("Mean without removing outliers:", round(m1, 2), "n")
  cat("Mean if we remove outliers:", round(m2, 2), "n")
  response <- readline(prompt="Do you want to remove outliers and to replace with NA? [yes/no]: ")
  if(response == "y" | response == "yes"){
    dt[as.character(substitute(var))] <- invisible(var_name)
    assign(as.character(as.list(match.call())$dt), dt, envir = .GlobalEnv)
    cat("Outliers successfully removed", "n")
    return(invisible(dt))
  } else{
    cat("Nothing changed", "n")
    return(invisible(var_name))
  }
}

```
Build a function to calculate percentage of missing values in Columns and Rows

```{r}
pMiss = function(x){
  sum(is.na(x))/length(x)*100
}
```

```{r}
par(mfrow=c(1, 1))
boxplot(Data$Order.Quantity)
summary(boxplot(Data$Order.Quantity))
boxplot.stats(Data$Order.Quantity)$out 
outlierKD(Data, Order.Quantity)

```

```{r}
aggregate(Data$Order.Quantity,by=list(Category=Data$Product.Container), FUN=sum)
aggregate(Data$Order.Quantity,by=list(Category=Data$Product.Sub.Category), FUN=sum)
aggregate(Data$Order.Quantity,by=list(Category=Data$Ship.Mode), FUN=sum)

aggregate(Data$Order.Quantity,by=list(Category=Data$Product.Container), FUN=mean)
aggregate(Data$Order.Quantity,by=list(Category=Data$Product.Sub.Category), FUN=mean)
aggregate(Data$Order.Quantity,by=list(Category=Data$Ship.Mode), FUN=mean)

```
```{r}
boxplot(Data$Sales)
summary(boxplot(Data$Sales))
aggregate(Data$Sales,by=list(Category=Data$Product.Container), FUN=sum)
aggregate(Data$Sales,by=list(Category=Data$Product.Sub.Category), FUN=sum)
aggregate(Data$Sales,by=list(Category=Data$Ship.Mode), FUN=sum)

aggregate(Data$Sales,by=list(Category=Data$Product.Container), FUN=mean)
aggregate(Data$Sales,by=list(Category=Data$Product.Sub.Category), FUN=mean)
aggregate(Data$Sales,by=list(Category=Data$Ship.Mode), FUN=mean)
```

```{r}

boxplot.stats(Data$Sales)$out 
outlierKD(Data, Sales)

```
Correlation plot

```{r}
my_num_data <- Data[, sapply(Data, is.numeric)]
cor.Data<-cor(my_num_data)
round(cor.Data,2)
library(corrplot)
corrplot(cor.Data,method="circle",bg="white")

```
Find Percentage of missing values in each column
```{r}
col_miss = apply(Data,2,pMiss)  ## 2 is for Columns
col_miss

```
Understanding of data balancing nature
```{r}
table(Data$Ship.Mode)
print(prop.table(table(Data$Ship.Mode)))

```
```{r}
attach(Data)
names(Data)
par(mfrow=c(1, 1))

interaction.plot(Product.Container,Ship.Mode,Sales,
                 col = c("Red","Blue","Green"))

interaction.plot(Product.Container,Ship.Mode,Order.Quantity,
                 col = c("Red","Blue","Green"))

interaction.plot(Product.Sub.Category,Ship.Mode,Sales,
                 col = c("Red","Blue","Green"))

interaction.plot(Product.Sub.Category,Ship.Mode,Order.Quantity,
                 col = c("Red","Blue","Green"))

```

```{r}
Data %>%
  keep(is.numeric) %>% 
  gather() %>% 
  ggplot(aes(value)) +
  facet_wrap(~ key, scales = "free") +
  geom_histogram()

## density diagram for all the variables
Data %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_density()   

## density diagram for all the variables
Data %>%
  keep(is.numeric) %>%                     # Keep only numeric columns
  gather() %>%                             # Convert to key-value pairs
  ggplot(aes(value)) +                     # Plot the values
  facet_wrap(~ key, scales = "free") +   # In separate panels
  geom_bar() 

```

```{r}
table(Ship.Mode)
prop.table(table(Ship.Mode))
```

```{r}
library(plotrix)
pie3D(prop.table(table(Ship.Mode)), 
      main='Ship mode', 
      #explode=0.05, 
      labels=c("Delivery Truck","Express Air", "Regular Air"), 
      col = c("Turquoise", "Medium Sea Green","Red") 
)
```
```{r}
table(Product.Container)
prop.table(table(Product.Container))
```

```{r}
pie3D(prop.table(table(Product.Container)), 
      main='Product Container', 
      #explode=0.05, 
      labels=c("Jumbo Box", "Jumbo Drum","Large Box","Medium Box","Small Box","Small Pack","Wrap Bag"), 
      col = c("Red", "Blue","Green","Yellow","Orange","Brown","Turquoise") 
)
```
```{r}
vis_summary <- ggplot(Data, aes(x = Data$Order.Quantity, y = Data$Sales)) +
  facet_grid(~ Data$Product.Container + Data$Ship.Mode)+
  geom_boxplot(na.rm = TRUE, colour = "#3366FF",outlier.colour = "red", outlier.shape = 1) +
  labs(x = "Order Quantity", y = "Sales") +
  scale_x_continuous() +
  scale_y_continuous() +
  theme(legend.position="bottom", legend.direction="horizontal")
vis_summary
```
Model Building

```{r}
#install.packages("caret",dependencies = TRUE)
library(caret)

inventory_df <- Data

colnames(inventory_df)

inventory_df$Order.Date <- NULL
inventory_df$Order.ID <- NULL
inventory_df$Product.Name <- NULL

# Data partition
intrain <- createDataPartition(y = inventory_df$Ship.Mode, p = 0.70, list = FALSE)
train_df <- inventory_df[intrain,]
test_df <- inventory_df[-intrain,]

dim(train_df)
dim(test_df)

```

##### Multi-Nominal Logistic Regression

```{r}
mlr.fit  <- multinom(train_df$Ship.Mode ~ ., train_df, maxit =1000)

summary(mlr.fit)

# Prediction on Train_data
mlr_train_pred <- predict(mlr.fit, train_df, type = "class")

# Accuracy and other metrics on train_data
confusionMatrix(mlr_train_pred, train_df$Ship.Mode)

# Predictions on test data
mlr_test_pred <- predict(mlr.fit, test_df, type = "class")

# Accuracy and other metrics on test data
confusionMatrix(mlr_test_pred, test_df$Ship.Mode)

```
# Random Forest

```{r}
set.seed(786)
# Training with Random forest model
modfit.rf <- randomForest(Ship.Mode ~. , data=train_df)

# Predict the training set with the trained model
rf_train_pred <- predict(modfit.rf, train_df, type = "class")

# Accuracy and other metrics on train data
confusionMatrix(rf_train_pred, train_df$Ship.Mode)

# Predict the test set with the trained model
rf_test_pred <- predict(modfit.rf, test_df, type = "class")

# Accuracy and other metrics on train data
confusionMatrix(rf_test_pred, test_df$Ship.Mode)

```
CART

```{r}
r.ctrl <- rpart.control(minsplit = 100,
                        minbucket = 10,
                        cp = 0,
                        xval = 10
)
modfit.cart <- rpart(formula = Ship.Mode~.,
            data = train_df,
            method = "class",
            control = r.ctrl)

library(rattle) 
library(RColorBrewer) 
fancyRpartPlot(modfit.cart, 
               uniform = TRUE, 
               main = "Final Tree", 
               palettes = c("Blues", "Oranges"))


# Predict the training set with the trained model
cart_train_pred <- predict(modfit.cart, train_df, type = "class")

# Accuracy and other metrics on train data
confusionMatrix(cart_train_pred, train_df$Ship.Mode)

# Predict the test set with the trained model
cart_test_pred <- predict(modfit.cart, test_df, type = "class")

# Accuracy and other metrics on train data
confusionMatrix(cart_test_pred, test_df$Ship.Mode)

```

4. Model Performance Measures


Observation: Decision tress performs better in terms of accuracy of overall model. The stability in train & test dataset is equal and model performs better. Random Forest has highest accuracy in train dataset but in test it is dropping by ~10%


Observation: Decision tress performs better in terms of accuracy of overall model. The stability in train & test dataset is equal and model performs better. Random Forest has highest accuracy in train dataset but in test it is dropping by ~10%
5. Business Interpretation & Recommendation
1.	The dataset is about inventory of a firm. The data consists of 7,853 rows with 8 columns. The data spans from 2007 to 2010 timeframe where the business seems to be steadily growing. The objective of the exercise is to find the optimal ship mode for the order placed by the customers.
2.	~ 75% of the orders were placed by ‘Regular Air’ shipping mode which also suggests that customers are more inclined towards Regular Air travel shipping mode. The least preferred is Express Air shipping mode. The reason could be high charges levied towards Express Air travel which takes lesser time to deliver the orders.
3.	The above boxplots Ship.Mode Vs Sales & Order.Quantity shows following observations
a)	Sales numbers are higher when using Delivery_Truck as the shipping mode. It may be due to Delivery truck can be used to deliver large quantities of the orders
b)	The median order quantity for all the 3 modes of Ship_mode. It suggests that irrespective of any shipping mode the median order quantity remains same.
4.	Regular Air Shipping mode has the highest order followed by Delivery Truck. Express Air Shipping mode has lowest orders. It may be due to higher charges levied on Express modes.
5.	Paper has been purchased more followed by Binders , Telephones & office furnishings
6.	Global High-Back Leather has been purchased more followed by Conference Tables & Doorstop
7.	Office Machines are purchased more followed by Tables, Telephones & Communication
8.	Regular Air has highest sales followed by Delivery Truck. 
9.	Thursday has the higher order quantity across 2007-2010.
10.	Sales are increasing YoY. Wednesday has higher average sales across the Year. Weekend sales are higher when compared to other days
11.	Decision Trees perform better in terms of accuracy & stability on train & test dataset. 
12.	The Business teams can utilize the Decision Tree model for performing shipping mode for orders based on historical data. This will reduce cost incurred on the shipping modes.
